---
title: "EDS230: Assignment 7"
author: "Sydney Rilum"
date: "5/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(deSolve)
library(sensitivity)
```

Consider the following model of forest growth (where forest size in measured in units of carbon (C))

dC/dt=r∗C for forests where C is below a threshold canopy closure

dC/dt=g∗(1−C/K) for forests where carbon is at or above the threshold canopy closure

and K is a carrying capacity in units of carbon

The size of the forest (C), Canopy closure threshold and carrying capacity are all in units of carbon

You could think of the canopy closure threshold as the size of the forest at which growth rates change from exponential to linear

You can think of r, as early exponential growth rate and g as the linear growth rate once canopy closure has been reached

## Here’s what you will do

Implement this model in R (as a differential equation) (Grading 25% - we will look for correct implementation but also good programming style - meaningful variable names, comments/documentation throughout )

```{r}
# source function
source("dforest.R")
```

Run the model for 300 years (using the ODE solver) starting with an initial forest size of 10 kg/C, and using the following parameters

canopy closure threshold of 50 kgC

K = 250 kg C (carrying capacity)

r= 0.01 (exponential growth rate before before canopy closure)

g = 2 kg/year (linear growth rate after canopy closure)

```{r}
# set parameters
r = 0.01
K = 250
g = 2

cct = 50
C = 10

# gets results for 300 years (evaluating every year)
nyears = seq(from = 1, to = 300)
parameters = list(r = r,
                  K = K,
                  g = g)

# use ODE solver to run forest growth function
results = ode(y = C, times = nyears, func = dforest, cct = cct, parms = parameters) 

colnames(results) = c("year", "C")

# turn it into a data frame
results = as.data.frame(results)
```


Graph the results. (15% - we look for an appropriate graphs and good visualization practice - labels, etc)

```{r}
# graph model results 
ggplot(results, aes(year, C)) +
  geom_line() +
  labs(y = "Forest Size (kg C)",
       title = "Forest Growth Model")
```

Run a sobol sensitivity analysis that explores how the estimated maximum and mean forest size (e.g maximum and mean values of C over the 300 years) varies with the pre canopy closure growth rate (r) and post-canopy closure growth rate (g) and canopy closure threshold and carrying capacity(K)

Assume that parameters are all normally distributed with means as given above and standard deviation of 10% of mean value

```{r}
# setting parameter estimates for sensitivity analysis
nyears = 300
r = rnorm(mean = 0.01, sd = 0.01*0.1, n = nyears)
K = rnorm(mean = 250, sd = 250*0.1, n = nyears)
g = rnorm(mean = 2, sd = 2*0.1, n = nyears)

# bind parameters to data frame
X1 = cbind.data.frame(r = r, K = K, g = g)

# repeat for second random sample
r = rnorm(mean = 0.01, sd = 0.01*0.1, n = nyears)
K = rnorm(mean = 250, sd = 250*0.1, n = nyears)
g = rnorm(mean = 2, sd = 2*0.1, n = nyears)

# bind parameters to data frame
X2 = cbind.data.frame(r = r, K = K, g = g)

# create sobel object and get sets of parameters for running the model
sens_P = sobolSalt(model = NULL, X1, X2, nboot = 300)

# lets add names 
colnames(sens_P$X) = c("r","K","g")

# our parameter sets are
head(sens_P$X)
```

```{r}
# run our differential equation and keep the output
# BUT
# what output do we want  to keep
# how about maximum population if we run the model for 200 years, and how many years to get to the carrying capacity

# for illustration lets look at running just one parameter sets and summarizing results
sens_P$X[1,]
# recall ODE needs ALL of our parameters in a single list 
# initial population and timnes for which we want output 
C = 10

# gets results for 200 years (evaluating every year)
nyears = seq(from = 1, to = 300)
parmaters = list(r=sens_P$X[1,"r"], 
                 K = sens_P$X[1,"K"],
                 g = sens_P$X[1,"g"])

result = ode(y= C, times = nyears, func = dforest, cct = cct, parms = parameters)

head(result)
colnames(result)=c("time","C")

# turn it into a data frame
result = as.data.frame(result)

ggplot(result, aes(time, C)) + 
  geom_point()
```



Graph the results of the sensitivity analysis as a box plot of maximum forest size and a plot of the two Sobol indices (S and T). (25% - correct implementation of Sobol and good graphing style -label etc )

```{r}

```


In 2-3 sentences, discuss what the results of your simulation might mean for climate change impacts on forest growth (e.g think about what parameters climate change might influence ). (25% - wee look for reasonable discussion that uses the results from your analysis and give extra points for discussions that offer particularly creative or insightful commentary)



Submit R markdown with model implementation, graphs and sensitivity analysis and R file with your model

(Final 10% for well organized clear R markdown)

Work in Groups - Due Thursday May 19th